---
title: LCMS
---

## Setup

```{r}
library(data.table)
library(ggplot2)
library(here)
library(RColorBrewer)
```


## Donor 3, AS0013B, Midazolam
### Peak status'
Check peak status' for any amounts below what below/above the limit of our standard curve.
```{r}
# Load data
donor_3_peak_status <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0013_peak_status.rds"))

donor_3_peak_status
```
**Should this be omitted? It is below the lowest point on our standard curve, but omitting it would make the mean higher, than it really is? For now it is included in calculated means.**


### **Plots**
```{r}
# Load processed data
donor_3_data <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0013.rds"))


# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```

IL-6 effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_6_AS0013

plot_data <- 
  donor_3_data[time_incubation == 0.5 
               & treatment %in% group_IL6
               ]


act_3A4_IL_6_AS0013 <- 
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7,
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
    ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-6: Donor 3",
          subtitle = "Facetted on treatment duration in hours"
          ) +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
        )


act_3A4_IL_6_AS0013

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_6_AS0013.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

IL-1B effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_1B_AS0013


plot_data <- 
  donor_3_data[time_incubation == 0.5 &
                 treatment %in% group_IL1B
               ]

act_3A4_IL_1B_AS0013 <-
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
  ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-1B: Donor 3",
          subtitle = "Facetted on treatment duration in hours") +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
  )

act_3A4_IL_1B_AS0013

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_1B_AS0013.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

## Donor 2, AS0008, midazolam
Load data
Check peak status' for any amounts below what below/above the limit of our standard curve.
```{r}
donor_2_peak_status <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0008_peak_status.rds"))

# Check peak status'
donor_2_peak_status |> knitr::kable()
```
Samples 52, 153, 162, 163, 172, and 173 had no baselcocktail added to there wells, and that is why there is no peaks. Sample 171 did not have a spheroid in the well. These samples are removed from the datatable before calculating means. However, there are two samples left with "response low", which are included in calculating means. 



### **Plots**
```{r}
# Load processed data
donor_2_data <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0008.rds"))


# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```

IL-6 effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_6_AS0008

plot_data <- 
  donor_2_data[time_incubation == 0.5 
               & treatment %in% group_IL6
               ]


act_3A4_IL_6_AS0008 <- 
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7,
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
    ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-6: Donor 2",
          subtitle = "Facetted on treatment duration in hours"
          ) +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
        )


act_3A4_IL_6_AS0008

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_6_AS0008.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

IL-1B effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_1B_AS0008


plot_data <- 
  donor_2_data[time_incubation == 0.5 &
                 treatment %in% group_IL1B
               ]

act_3A4_IL_1B_AS0008 <-
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
  ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-1B: Donor 2",
          subtitle = "Facetted on treatment duration in hours") +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
  )

act_3A4_IL_1B_AS0008

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_1B_AS0008.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```


## Main plots

load data

```{r}
data <- readRDS(here("notebooks/LCMS/data_processed/final_data.rds"))
```

```{r}
# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```


### Facetted on time 
IL-6
```{r}
#| label: fig-act-cyp3A4-IL6
#| fig-cap: "24 hour timepoints removed"
#| fig-width: 10
#| fig-height: 5

plot_data_IL6 <- data[time_incubation == 0.5 
                      & treatment %in% group_IL6
                      & time_treatment != 24]


act_3A4_IL6 <- ggplot(data = plot_data_IL6) +
  geom_point(
    mapping = aes(x = treatment, y = mean_relative_amount, color = factor(time_treatment)),
    size = 3.5
  ) +
  stat_summary(
    fun = "mean",
    geom = "point",
    mapping = aes(x = treatment, y = mean_relative_amount),
    size = 5
  ) +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(~ time_treatment, ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-6 treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration") + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  geom_hline(yintercept = 1, linetype = "dashed")



act_3A4_IL6
```
CONSIDER REMOVING THE OUTLIER IN 120 hour IL-6 and IL1B. See the individual plots above.



IL-1B
```{r}
#| label: fig-act-cyp3A4-IL1B
#| fig-cap: "24 hour timepoints removed"
#| fig-width: 10
#| fig-height: 5

plot_data_I1B <- data[time_incubation == 0.5 
                      & treatment %in% group_IL1B
                      & time_treatment != 24]


ggplot(data = plot_data_I1B) +
  geom_point(
    mapping = aes(x = treatment, y = mean_relative_amount, color = factor(time_treatment)),
    size = 3.5
  ) +
  stat_summary(
    fun = "mean",
    geom = "point",
    mapping = aes(x = treatment, y = mean_relative_amount),
    size = 5
  ) +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(facets = vars(time_treatment)) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-1B treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration"
       ) + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  geom_hline(yintercept = 1, linetype = "dashed")


```

### Timepoints combined
IL-6
```{r}
#| label: fig-act-il6-combined
#| fig-cap: "Overall mean CYP3A4 activity relative to vehicle control after IL-6 treatment. 24 hour timepoints removed"

IL6_combined_data <- 
  data[time_incubation == 0.5
       & treatment %in% group_IL6
       & time_treatment != 24]

IL6_combined_data[,
                  overall_mean := mean(mean_relative_amount),
                  by = .(treatment)]


ggplot(data = IL6_combined_data) +
  geom_point(aes(x = treatment,
                 y = relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 5,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = overall_mean),
             size = 10,
             alpha = 0.1
             ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-6 treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration"
       ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  theme(#axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  scale_y_continuous(limits = c(0, NA))
```


IL-1B
```{r}
#| label: fig-act-il1b-combined
#| fig-cap: "Overall mean CYP3A4 activity relative to vehicle control after IL-1B treatment. 24 hour timepoints removed"

IL1B_combined_data <- 
  data[time_incubation == 0.5
       & treatment %in% group_IL1B
       & time_treatment != 24]

IL1B_combined_data[,
                  overall_mean := mean(mean_relative_amount),
                  by = .(treatment)]


ggplot(data = IL1B_combined_data) +
  geom_point(aes(x = treatment,
                 y = relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 5,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = overall_mean),
             size = 10,
             alpha = 0.1
             ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-1B treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration"
       ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  theme(#axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  scale_y_continuous(limits = c(0, NA))
```



## EC50








