---
title: LCMS
---

## Setup

```{r}
library(data.table)
library(ggplot2)
library(here)
library(RColorBrewer)
```


## Donor 3, AS0013B, Midazolam
### Peak status'
Check peak status' for any amounts below what below/above the limit of our standard curve.
```{r}
# Load data
donor_3_peak_status <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0013_peak_status.rds"))

donor_3_peak_status
```
**Should this be omitted? It is below the lowest point on our standard curve, but omitting it would make the mean higher, than it really is? For now it is included in calculated means.**


### **Plots**
```{r}
# Load processed data
donor_3_data <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0013.rds"))


# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```

IL-6 effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_6_AS0013
#| fig-cap: "Donor 3: Activity, IL-6"

plot_data <- 
  donor_3_data[time_incubation == 0.5 
               & treatment %in% group_IL6
               ]


act_3A4_IL_6_AS0013 <- 
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7,
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
    ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-6: Donor 3",
          subtitle = "Facetted on treatment duration in hours"
          ) +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
        )


act_3A4_IL_6_AS0013

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_6_AS0013.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

IL-1B effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_1B_AS0013
#| fig-cap: "Donor 3: Activity, IL-1B"


plot_data <- 
  donor_3_data[time_incubation == 0.5 &
                 treatment %in% group_IL1B
               ]

act_3A4_IL_1B_AS0013 <-
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
  ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-1B: Donor 3",
          subtitle = "Facetted on treatment duration in hours") +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
  )

act_3A4_IL_1B_AS0013

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_1B_AS0013.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

## Donor 2, AS0008, midazolam
Load data
Check peak status' for any amounts below what below/above the limit of our standard curve.
```{r}
donor_2_peak_status <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0008_peak_status.rds"))

# Check peak status'
donor_2_peak_status |> knitr::kable()
```
Samples 52, 153, 162, 163, 172, and 173 had no baselcocktail added to there wells, and that is why there is no peaks. Sample 171 did not have a spheroid in the well. These samples are removed from the datatable before calculating means. However, there are two samples left with "response low", which are included in calculating means. 



### **Plots**
```{r}
# Load processed data
donor_2_data <- 
  readRDS(here("notebooks/LCMS/data_processed/AS0008.rds"))


# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```

IL-6 effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_6_AS0008
#| fig-cap: "Donor 2: Activity, IL-6"

plot_data <- 
  donor_2_data[time_incubation == 0.5 
               & treatment %in% group_IL6
               ]


act_3A4_IL_6_AS0008 <- 
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7,
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
    ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-6: Donor 2",
          subtitle = "Facetted on treatment duration in hours"
          ) +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
        )


act_3A4_IL_6_AS0008

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_6_AS0008.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```

IL-1B effect on CYP3A4
```{r}
#| label: fig-act_3A4_IL_1B_AS0008
#| fig-cap: "Donor 2: Activity, IL-1B"


plot_data <- 
  donor_2_data[time_incubation == 0.5 &
                 treatment %in% group_IL1B
               ]

act_3A4_IL_1B_AS0008 <-
  ggplot(data = plot_data) +
  geom_point(aes(x = treatment,
                 y = mean_relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 8,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = relative_amount
                 ),
             size = 4,
             alpha = 0.3
             ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(
    name = "OH-midazolam formation relative to vehicle control", 
    breaks = seq(from = 0, to = 3, by= 0.25)
    ) +
  facet_wrap(
    facets = vars(factor(time_treatment))
  ) +
  theme_bw() +
  xlab("Treatment") + 
  ggtitle("CYP3A4 activity after treatment with IL-1B: Donor 2",
          subtitle = "Facetted on treatment duration in hours") +
  ylab("Metabolite formation relative to vehicle control") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.title.x = element_text(hjust = 0.5, vjust = 0, face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none"
  )

act_3A4_IL_1B_AS0008

# ggsave(
#   plot = last_plot(),
#   filename = "3A4_IL_1B_AS0008.png",
#   device = "png",
#   path = here("notebooks/LCMS/output")
#   )
```


## Main plots

load data

```{r}
data <- readRDS(here("notebooks/LCMS/data_processed/final_data.rds"))
```

```{r}
# Create groups/look up tables for subsetting data
group_IL6 <- c(#"control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c(#"control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )
```


### Facetted on time 
IL-6
```{r}
#| label: fig-act-cyp3A4-IL6
#| fig-cap: "Mean CYP3A4 activity relative to vehicle control after IL-6 treatment. Facetted on time. 24 hour timepoints removed"
#| fig-width: 10
#| fig-height: 5

plot_data_IL6 <- data[time_incubation == 0.5 
                      & treatment %in% group_IL6
                      & time_treatment != 24]


act_3A4_IL6 <- ggplot(data = plot_data_IL6) +
  geom_point(
    mapping = aes(x = treatment, y = mean_relative_amount, color = factor(time_treatment)),
    size = 3.5
  ) +
  stat_summary(
    fun = "mean",
    geom = "point",
    mapping = aes(x = treatment, y = mean_relative_amount),
    size = 5
  ) +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(~ time_treatment, ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-6 treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration") + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  geom_hline(yintercept = 1, linetype = "dashed")



act_3A4_IL6
```
CONSIDER REMOVING THE OUTLIER IN 120 hour IL-6 and IL1B. See the individual plots above.


IL-1B
```{r}
#| label: fig-act-cyp3A4-IL1B
#| fig-cap: "Mean CYP3A4 activity relative to vehicle control after IL-1B treatment. Facetted on time. 24 hour timepoints removed"
#| fig-width: 10
#| fig-height: 5

plot_data_I1B <- data[time_incubation == 0.5 
                      & treatment %in% group_IL1B
                      & time_treatment != 24]


act_3A4_IL1B <- 
  ggplot(data = plot_data_I1B) +
    geom_point(
      mapping = aes(x = treatment, y = mean_relative_amount, color = factor(time_treatment)),
      size = 3.5
    ) +
    stat_summary(
      fun = "mean",
      geom = "point",
      mapping = aes(x = treatment, y = mean_relative_amount),
      size = 5
    ) +
    scale_color_brewer(palette = "Set2") +
    facet_wrap(facets = vars(time_treatment)) +
    labs(title = "CYP3A4 activity relative to vehicle control after IL-1B treatment",
         x = "Treatment",
         y = "Expression relative to vehicle control",
         color = "Treatment duration"
         ) + 
    scale_y_continuous(limits = c(0, NA)) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          axis.title.y = element_text(vjust = 3, face = "bold"),
          axis.title.x = element_text(vjust = -1.5, face ="bold"),
          plot.title = element_text(face = "bold"),
          legend.title = element_text(face = "bold"),
          legend.position = "bottom"
       ) +
    geom_hline(yintercept = 1, linetype = "dashed")

act_3A4_IL1B
```

### Timepoints combined
IL-6
```{r}
#| label: fig-act-il6-combined
#| fig-cap: "Overall mean CYP3A4 activity relative to vehicle control after IL-6 treatment. 24 hour timepoints removed"

IL6_combined_data <- 
  data[time_incubation == 0.5
       & treatment %in% group_IL6
       & time_treatment != 24]

IL6_combined_data[,
                  overall_mean := mean(mean_relative_amount),
                  by = .(treatment)]


ggplot(data = IL6_combined_data) +
  geom_point(aes(x = treatment,
                 y = relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 5,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = overall_mean),
             size = 10,
             alpha = 0.1
             ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-6 treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration"
       ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  theme(#axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  scale_y_continuous(limits = c(0, NA))
```


IL-1B
```{r}
#| label: fig-act-il1b-combined
#| fig-cap: "Overall mean CYP3A4 activity relative to vehicle control after IL-1B treatment. 24 hour timepoints removed"

IL1B_combined_data <- 
  data[time_incubation == 0.5
       & treatment %in% group_IL1B
       & time_treatment != 24]

IL1B_combined_data[,
                  overall_mean := mean(mean_relative_amount),
                  by = .(treatment)]


ggplot(data = IL1B_combined_data) +
  geom_point(aes(x = treatment,
                 y = relative_amount,
                 color = factor(time_treatment)
                 ),
             size = 5,
             alpha = 0.7
             ) +
  scale_color_brewer(palette = "Set2") +
  geom_point(aes(x = treatment,
                 y = overall_mean),
             size = 10,
             alpha = 0.1
             ) +
  labs(title = "CYP3A4 activity relative to vehicle control after IL-1B treatment",
       x = "Treatment",
       y = "Expression relative to vehicle control",
       color = "Treatment duration"
       ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  theme(#axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.y = element_text(vjust = 3, face = "bold"),
        axis.title.x = element_text(vjust = -1.5, face ="bold"),
        plot.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom"
     ) +
  scale_y_continuous(limits = c(0, NA))
```



## EC50


```{r}
EC50_data <- readRDS(here("notebooks/LCMS/data_processed/final_data.rds"))

# Create groups/look up tables for subsetting data
group_IL6 <- c("control", 
               #"medium", 
               "0.01 ng/ml IL-6", 
               "0.1 ng/ml IL-6",
               "1 ng/ml IL-6", 
               "10 ng/ml IL-6"
               )

group_IL1B <- c("control", 
                #"medium", 
                "0.01 ng/ml IL-1B", 
                "0.1 ng/ml IL-1B",
                "1 ng/ml IL-1B",
                "10 ng/ml IL-1B"
                )

```

```{r}
EC50_IL6 <- EC50_data[treatment %in% group_IL6]

#fwrite(EC50_IL6, file = here("notebooks/LCMS/EC50_IL6.csv"))
```




```{r}
library(tidyverse)
library(minpack.lm)
```

```{r}
EC50_IL6 <- EC50_IL6 %>%
  mutate(Concentration = case_when(
    treatment == "control" ~ 0, # Assign 0 to control
    TRUE ~ as.numeric(gsub(".*?(\\d+\\.?\\d*).*", "\\1", treatment)) # Extract numeric values
  ))

# Check for NAs in Concentration to ensure parsing was successful
sum(is.na(EC50_IL6$Concentration))

```

```{r}

# The 4-parameter logistic model function remains the same
fourPL <- function(x, Top, Bottom, IC50, HillSlope) {
  Bottom + (Top - Bottom) / (1 + (x / IC50)^(-HillSlope))
}

# Ensure EC50_IL6 is filtered and contains no NA values for the fitting
EC50_IL6_filtered <- EC50_IL6 %>% filter(!is.na(Concentration) & Concentration > 0 & !is.na(relative_amount))

# Adjust initial parameter estimates
Top_est <- max(EC50_IL6_filtered$relative_amount, na.rm = TRUE)
Bottom_est <- min(EC50_IL6_filtered$relative_amount, na.rm = TRUE)
IC50_est <- median(EC50_IL6_filtered$Concentration, na.rm = TRUE) # Assuming this is a reasonable estimate
HillSlope_est <- 1 # Starting with a neutral slope

# Fit the model with adjusted initial estimates
model <- tryCatch({
  nlsLM(relative_amount ~ fourPL(Concentration, Top, Bottom, IC50, HillSlope), 
        data = EC50_IL6_filtered, 
        start = list(Top = Top_est, Bottom = Bottom_est, IC50 = IC50_est, HillSlope = HillSlope_est))
}, error = function(e) e)

# Check if model fitting was successful
if(inherits(model, "error")) {
  print("Model fitting encountered an error:")
  print(model$message)
} else {
  print("Model fitted successfully.")
}

```

```{r}
#| label: fig-dose-response-act-il6


# Generate predictions from the model across a range of concentrations
new_data <- data.frame(Concentration = seq(min(EC50_IL6_filtered$Concentration, na.rm = TRUE), 
                                           max(EC50_IL6_filtered$Concentration, na.rm = TRUE), length.out = 100))
new_data$Predicted <- predict(model, newdata = new_data)

# Plot the dose-response curve
ggplot() +
  geom_point(data = EC50_IL6_filtered, aes(x = Concentration, y = relative_amount), color = "blue") +
  geom_line(data = new_data, aes(x = Concentration, y = Predicted), color = "red") +
  labs(title = "Dose-response curve", x = "Concentration", y = "Relative Activity") +
  theme_minimal() +
  scale_x_log10()
```



```{r}
# Print the summary of the fitted model to see parameter estimates
summary(model)
```



