---
title: "Seqeuncing"
editor_options: 
  chunk_output_type: console
---
# Volcano plots

Setup
```{r}
# Always load same package versions
groundhog.day <- "2024-01-01"
groundhog::meta.groundhog(groundhog.day) 
pkgs <-  c("data.table",
           "readxl",
           "ggplot2",
           "ggrepel",
           "here")

groundhog.library(pkgs, groundhog.day)
```


## IL-6 pooled
No sig up or down regulated genes when pooling IL-6 10 and 0.1 ng/ml
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL6_vs_vehicle.xlsx")
il6_pooled_data <- read_xlsx(file, sheet = 1)


setDT(il6_pooled_data)
il6_pooled_data[logCPM > 1 & logFC > 1 & FDR < 0.05]
il6_pooled_data[logCPM > 1 & logFC < -1 & FDR < 0.05]
```
As seen in empty dataframes there are no sig up or downregulated.
Also seen in the volcano plot


```{r}
# Create a new column for labeling differential gene expression
il6_pooled_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il6_pooled_data$diffexpressed[il6_pooled_data$logFC > 1 & il6_pooled_data$FDR < 0.05 & il6_pooled_data$logCPM > 1] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il6_pooled_data$diffexpressed[il6_pooled_data$logFC < -1 & il6_pooled_data$FDR < 0.05 & il6_pooled_data$logCPM > 1] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il6_pooled_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il6_pooled_data$delabel[il6_pooled_data$diffexpressed != "NO"] <- il6_pooled_data$genes[il6_pooled_data$diffexpressed != "NO"]
```

```{r}
# Create the volcano plot using ggplot2
plot_volcano_il6_pooled <- 
  ggplot(data=il6_pooled_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 7) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    #scale_x_continuous(breaks = seq(-11, 9, 2)) + # Customize x-axis breaks
    #scale_y_continuous(breaks = seq(0, 90, 10)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(), panel.grid.minor.x = element_blank())+ # Adjust minor grid lines
    ggtitle('Volcano Plot: IL6 vs Vehicle') # Add a title to the plot

plot_volcano_il6_pooled

```

## IL-6 10 ng/ml
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL6_10_vs_vehicle.xlsx")
il6_10_data <- read_xlsx(file, sheet = 1)
```


```{r}
#| label: fig-volcano-il6-10
#| fig-cap: "Volcano plot. IL-6 10 ng/ml vs. vehicle control. Vertical lines for fold change thresholds"
#| warning: false

# Create a new column for labeling differential gene expression
il6_10_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il6_10_data$diffexpressed[il6_10_data$logFC > 1 & il6_10_data$FDR < 0.05 & il6_10_data$logCPM > 1] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il6_10_data$diffexpressed[il6_10_data$logFC < -1 & il6_10_data$FDR < 0.05 & il6_10_data$logCPM > 1] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il6_10_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il6_10_data$delabel[il6_10_data$diffexpressed != "NO"] <- il6_10_data$genes[il6_10_data$diffexpressed != "NO"]

# Create the volcano plot using ggplot2
plot_volcano_il6_10 <- 
  ggplot(data=il6_10_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 10) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
    scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
          panel.grid.minor.x = element_blank(),# Remove minor grid lines
          legend.position = "bottom" 
          )+ 
    ggtitle('Volcano Plot: IL6 10 ng/ml vs Vehicle') # Add a title to the plot


plot_volcano_il6_10

# ggsave(filename = here("notebooks/seq/output/IL_6_10ng_volcano.png"))
```

Trying to ad a box around the out of bounds data points. They have a smaller FDR than can be handled.
```{r}
# # Adjusting the dataset to include points with FDR of 0
# il6_10_data$adjusted_FDR <- ifelse(il6_10_data$FDR == 0, 340, -log10(il6_10_data$FDR))
# 
# # Create the volcano plot using ggplot2
# plot_volcano_il6_10 <- 
#   ggplot(data=il6_10_data, aes(x=logFC, y=adjusted_FDR, col=diffexpressed, label=delabel)) + 
#     theme_minimal() +
#     geom_text_repel(max.overlaps = 10) +
#     scale_color_manual(values=c("blue", "black", "red")) +
#     geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") +
#     geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Use the new adjusted value
#     geom_point() +
#     labs(color = 'Direction of gene expression',
#          x = expression("log"[2]*"FC (IL6 vs Vehicle)"),
#          y = expression("-log"[10]*"Adjusted FDR")) +
#     scale_x_continuous(limits = c(-8, 10)) +
#     scale_y_continuous(limits = c(0, 450)) +
#     theme(
#       panel.grid.minor.y = element_blank(),
#       panel.grid.minor.x = element_blank(),
#       legend.position = "bottom"
#     ) +
#     annotate("rect", xmin = 1, xmax = 10, ymin = 320, ymax = 340, alpha = .2, color = "black") +
#     ggtitle('Volcano Plot: IL6 10 ng/ml vs Vehicle')
# 
# # Display the plot
# plot_volcano_il6_10


```

## IL-6 0.1 ng/ml
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL6_0.1_vs_vehicle.xlsx")
il6_0.1_data <- read_xlsx(file, sheet = 1)


setDT(il6_0.1_data)
il6_0.1_data[logCPM > 1 & FDR < 0.05 & logFC > 0] |> nrow()
il6_0.1_data[logCPM > 1 & FDR < 0.05 & logFC < 0] |> nrow()
```
Ikke nogen der er signifikant opreguleret når logFC > 1
Kun to signifikant nedregulerede ved logFC < -1

```{r}
#| label: fig-volcano-il6-0.1
#| fig-cap: "Volcano plot. IL-6 0.1 ng/ml vs. vehicle control. Vertical lines for fold change thresholds. Note few significant above 2 fold change"
#| warning: false

# Create a new column for labeling differential gene expression
il6_0.1_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il6_0.1_data$diffexpressed[il6_0.1_data$FDR < 0.05 & il6_0.1_data$logCPM > 1 & il6_0.1_data$logFC > 0] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il6_0.1_data$diffexpressed[il6_0.1_data$FDR < 0.05 & il6_0.1_data$logCPM > 1 & il6_0.1_data$logFC < 0] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il6_0.1_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il6_0.1_data$delabel[il6_0.1_data$diffexpressed != "NO"] <- il6_0.1_data$genes[il6_0.1_data$diffexpressed != "NO"]

# Create the volcano plot using ggplot2
plot_volcano_il6_0.1 <- 
  ggplot(data=il6_0.1_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 10) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    #scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
    #scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
          panel.grid.minor.x = element_blank(),# Remove minor grid lines
          legend.position = "bottom" 
          )+ 
    ggtitle('Volcano Plot: IL6 0.1 ng/ml vs Vehicle') # Add a title to the plot


plot_volcano_il6_0.1

# ggsave(filename = here("notebooks/seq/output/IL_6_01ng_volcano.png"))

```


## IL-1B pooled
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL1B_vs_vehicle.xlsx")
il1B_pooled_data <- read_xlsx(file, sheet = 1)


setDT(il1B_pooled_data)
il1B_pooled_data[logCPM > 1 & FDR < 0.05 & logFC > 1] |> nrow()
il1B_pooled_data[logCPM > 1 & FDR < 0.05 & logFC < 1] |> nrow()
```

```{r}
#| label: fig-volcano-il1B-pooled
#| fig-cap: "Volcano plot. IL-1B pooled vs. vehicle control. Vertical lines for fold change thresholds"
#| warning: false

# Create a new column for labeling differential gene expression
il1B_pooled_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il1B_pooled_data$diffexpressed[il1B_pooled_data$FDR < 0.05 & il1B_pooled_data$logCPM > 1 & il1B_pooled_data$logFC > 0] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il1B_pooled_data$diffexpressed[il1B_pooled_data$FDR < 0.05 & il1B_pooled_data$logCPM > 1 & il1B_pooled_data$logFC < 0] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il1B_pooled_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il1B_pooled_data$delabel[il1B_pooled_data$diffexpressed != "NO"] <- il1B_pooled_data$genes[il1B_pooled_data$diffexpressed != "NO"]

# Create the volcano plot using ggplot2
plot_volcano_il1B_pooled <- 
  ggplot(data=il1B_pooled_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 10) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    #scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
    #scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
          panel.grid.minor.x = element_blank(),# Remove minor grid lines
          legend.position = "bottom" 
          )+ 
    ggtitle('Volcano Plot: IL1B 10 ng/ml vs Vehicle') # Add a title to the plot


plot_volcano_il1B_pooled

# ggsave(filename = here("notebooks/seq/output/IL_1B_10ng_volcano.png"))

```


## IL-1B 10 ng/ml
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL1B_10_vs_vehicle.xlsx")
il1B_10_data <- read_xlsx(file, sheet = 1)


setDT(il1B_10_data)
il1B_10_data[logCPM > 1 & FDR < 0.05 & logFC > 1] |> nrow()
il1B_10_data[logCPM > 1 & FDR < 0.05 & logFC < 1] |> nrow()
```


```{r}
#| label: fig-volcano-il1B-10
#| fig-cap: "Volcano plot. IL-1B 10 ng/ml vs. vehicle control. Vertical lines for fold change thresholds"
#| warning: false

# Create a new column for labeling differential gene expression
il1B_10_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il1B_10_data$diffexpressed[il1B_10_data$FDR < 0.05 & il1B_10_data$logCPM > 1 & il1B_10_data$logFC > 0] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il1B_10_data$diffexpressed[il1B_10_data$FDR < 0.05 & il1B_10_data$logCPM > 1 & il1B_10_data$logFC < 0] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il1B_10_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il1B_10_data$delabel[il1B_10_data$diffexpressed != "NO"] <- il1B_10_data$genes[il1B_10_data$diffexpressed != "NO"]

# Create the volcano plot using ggplot2
plot_volcano_il1B_0.1 <- 
  ggplot(data=il1B_10_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 10) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    #scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
    #scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
          panel.grid.minor.x = element_blank(),# Remove minor grid lines
          legend.position = "bottom" 
          )+ 
    ggtitle('Volcano Plot: IL1B 10 ng/ml vs Vehicle') # Add a title to the plot


plot_volcano_il1B_0.1

# ggsave(filename = here("notebooks/seq/output/IL_1B_10ng_volcano.png"))

```


## IL-1B 0.1 ng/ml
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/IL1B_0.1_vs_vehicle.xlsx")
il1B_0.1_data <- read_xlsx(file, sheet = 1)


setDT(il1B_0.1_data)
il1B_0.1_data[logCPM > 1 & FDR < 0.05 & logFC > 1] |> nrow()
il1B_0.1_data[logCPM > 1 & FDR < 0.05 & logFC < 1] |> nrow()
```


```{r}
#| label: fig-volcano-il1B-0.1
#| fig-cap: "Volcano plot. IL-1B 0.1 ng/ml vs. vehicle control. Vertical lines for fold change thresholds"
#| warning: false

# Create a new column for labeling differential gene expression
il1B_0.1_data$diffexpressed <- "NO" # Default value for genes not differentially expressed

# Identify and label upregulated genes
# Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
il1B_0.1_data$diffexpressed[il1B_0.1_data$FDR < 0.05 & il1B_0.1_data$logCPM > 1 & il1B_0.1_data$logFC > 0] <- "UP"

# Identify and label downregulated genes
# Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
il1B_0.1_data$diffexpressed[il1B_0.1_data$FDR < 0.05 & il1B_0.1_data$logCPM > 1 & il1B_0.1_data$logFC < 0] <- "DOWN"

# Initialize 'delabel' column to NA for all rows to create the column
il1B_0.1_data$delabel <- NA

# Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
il1B_0.1_data$delabel[il1B_0.1_data$diffexpressed != "NO"] <- il1B_0.1_data$genes[il1B_0.1_data$diffexpressed != "NO"]

# Create the volcano plot using ggplot2
plot_volcano_il1B_0.1 <- 
  ggplot(data=il1B_0.1_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
    theme_minimal() + # Use a minimal theme for the plot
    geom_text_repel(max.overlaps = 7) + # Use repelling text to avoid label overlap
    scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
    geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
    geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
    geom_point() + # Add points for each gene
    labs(color = 'Direction of gene expression', # Customize legend title
         x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
         y = expression("-log"[10]*"FDR")) + # Customize y-axis label
    #scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
    #scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
    theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
          panel.grid.minor.x = element_blank(),# Remove minor grid lines
          legend.position = "bottom" 
          )+ 
    ggtitle('Volcano Plot: IL1B 0.1 ng/ml vs Vehicle') # Add a title to the plot


plot_volcano_il1B_0.1

# ggsave(filename = here("notebooks/seq/output/IL_1B_01ng_volcano.png"))
```

## Abemaciclib 5 µM 
```{r}
# Import raw data from the specified Excel file and sheet
file <- here("notebooks/seq/data_raw/abe_vs_vehicle.xlsx")
abe_data <- read_xlsx(file, sheet = 1)


setDT(abe_data)
abe_data[logCPM > 1 & FDR < 0.05 & logFC > 1] 
abe_data[logCPM > 1 & FDR < 0.05 & logFC < 1] 
```


```{r}
# #| label: fig-abemaceclib
# #| fig-cap: "Volcano plot. Abemaciclib vs. vehicle control. Vertical lines for fold change thresholds"
# #| warning: false
# 
# # Create a new column for labeling differential gene expression
# abe_data$diffexpressed <- "NO" # Default value for genes not differentially expressed
# 
# # Identify and label upregulated genes
# # Genes with log2FoldChange > 1, FDR < 0.05, and logCPM > 1 are labeled as "UP"
# abe_data$diffexpressed[abe_data$FDR < 0.05 & abe_data$logCPM > 1 & abe_data$logFC > 0] <- "UP"
# 
# # Identify and label downregulated genes
# # Genes with log2FoldChange < -1, FDR < 0.05, and logCPM > 1 are labeled as "DOWN"
# abe_data$diffexpressed[abe_data$FDR < 0.05 & abe_data$logCPM > 1 & abe_data$logFC < 0] <- "DOWN"
# 
# # Initialize 'delabel' column to NA for all rows to create the column
# abe_data$delabel <- NA
# 
# # Now, you can safely assign gene names to the 'delabel' column for differentially expressed genes
# abe_data$delabel[abe_data$diffexpressed != "NO"] <- abe_data$genes[abe_data$diffexpressed != "NO"]
# 
# # Create the volcano plot using ggplot2
# plot_volcano_abe <- 
#   ggplot(data=abe_data, aes(x=logFC, y=-log10(FDR), col=diffexpressed, label=delabel)) + 
#     theme_minimal() + # Use a minimal theme for the plot
#     geom_text_repel() + # Use repelling text to avoid label overlap
#     scale_color_manual(values=c("blue", "black", "red")) + # Custom colors for differential expression
#     geom_vline(xintercept=c(-1, 1), col="dark grey", linetype = "dashed") + # Vertical lines for fold change thresholds
#     geom_hline(yintercept=-log10(0.05), col="dark grey", linetype = "dashed") + # Horizontal line for FDR threshold
#     geom_point() + # Add points for each gene
#     labs(color = 'Direction of gene expression', # Customize legend title
#          x = expression("log"[2]*"FC (IL6 vs Vehicle)"), # Customize x-axis label
#          y = expression("-log"[10]*"FDR")) + # Customize y-axis label
#     #scale_x_continuous(limits = c(-8, 10)) + # Customize x-axis breaks
#     #scale_y_continuous(limits = c(0, 320)) + # Customize y-axis breaks
#     theme(panel.grid.minor.y = element_blank(),# Remove minor grid lines
#           panel.grid.minor.x = element_blank(),# Remove minor grid lines
#           legend.position = "bottom",
#           panel.background = element_rect(fill = "white")
#           )+ 
#   coord_flip() +
#     ggtitle('Volcano Plot: 5 µM abemaciclib vs Vehicle') # Add a title to the plot 
# 
# 
# 
# plot_volcano_abe
# 
# ggsave(filename = here("notebooks/seq/output/abe_volcano.png"))
```

